
This thesis is the result of part of my work carried out in the context of two Projects: the \textsc{Datalift} Project (ANR-10-CORD-009), and the \textsc{Apps4Europe} Project. 
There are many people I want to thank since they kindly supported me in so many different ways for the successful completion of this thesis.


I am indebted to my thesis advisor Dr. Rapha\"{e}l Troncy for giving me the opportunity for a PhD. at EURECOM / Telecom ParisTech. Throughout my PhD  he provided helpful ideas and encouraging support in this exciting domain of Semantic Web . 

I would like to thank my committee members, the reviewers Prof. XXX  and Dr. XXXX, and furthermore the examiners Dr. XX and Dr. XXX for their precious time, shared positive insight and guidance.

I would like to express my deepest appreciation for all the colleagues and partners of Datalift project for all the exchanges during the duration of the project. I would like to thank specially Bernard, Pierre-Yves, Nathalie, Laurent. It was a pleasure to work and exchange with them.

My sincere thanks are due to all the members of the W3C Government Linked Data Working Group, specially to Bernadette Hyland, Boris Villaz\'{o}n-Terrazas, Richard Cyganiak, Dave Reynolds and Phil Archer. We had many intensive moments of exchange and collaborations.

I would like also to show my sincerest gratitude to all my mentors at OEG/UPM (Madrid) for providing me first skills on ontology engineering: Pr. Asuncion G\'{o}mez and Dr. Oscar Corcho. Their enthusiasm and dedication to research on this field were truly inspiring.


My warmest thanks to my colleagues who supported me during my Ph.D. Precisely, I would like to thank Houda, Giuseppe, Jos\'{e} Luis, Vuk and Ahmad. Also, I thank all those working at EURECOM, they made my stay at EURECOM very pleasant.

I would like to express my deepest thanks to my wife, Ver\'{o}nica \'{A}lvarez Aguilar, who gave me all her patience and comprehension since the beginning of this adventure. 
I owe my family my profound gratitude because they
always supported and believed in me: specially my parents Genevieve Djifack and Prosper Tabondjou; and all my brothers and sisters: Stephanie, Judith, Arlette, Mathias, Alvine, Edwige, Yannick.

Lastly, special thanks to my friends for their unwavering friendship, moral and infinite support.


 

