
%{\bf \Large Acronyms}\\

\chapter*{List of Publications}

\section*{Book Chapter}
\label{sec:book}
\begin{itemize}
\item \underline{\textbf{Ghislain Atemezing}} and {R}apha{\"e}l {T}roncy: \textbf{{M}ultimedia metadata}. {I}n {E}ncyclopedia of {S}ocial {N}etwork {A}nalysis and {M}ining, {S}pringer {V}erlag, 2014, {ISBN}: 978-1461461692
\end{itemize}

\section*{Journal}
\label{sec:journal}
\begin{enumerate}
 \item \underline{\textbf{{G}hislain} {A}temezing}, {O}scar {C}orcho,  {D}aniel {G}arijo, {J}os{\'e} {M}ora, {M}ar{\'i}a {P}oveda-{V}illal{\'o}n, {P}ablo {R}ozas, {D}aniel {V}ila-{S}uero and {B}oris {V}illaz{\'o}n-{T}errazas: \textbf{{T}ransforming meteorological data into linked data}. In {S}emantic {W}eb journal, {S}pecial {I}ssue on {L}inked {D}ataset descriptions, 2012. {IOS} {P}ress.

 \item Mari Carmen Su\'{a}rez-Figueroa, \underline{\textbf{Ghislain Atemezing}} and Oscar Corcho: \textbf{The landscape of multimedia ontologies in the last decade}. In Multimedia Tools and Applications, Vol. 55, Num. 3, December 2011.
\end{enumerate}

\section*{Conferences and Workshops}\label{sec:conf}
\begin{enumerate}
 \item \underline{\textbf{Ghislain Atemezing}} and Rapha{\"e}l Troncy: \textbf{{T}owards a linked-data based visualization wizard}. In 5th {I}nternational {W}orkshop on {C}onsuming {L}inked {D}ata (COLD 2014), {O}ctober 2014, {R}iva del {G}arda, {I}taly .

 \item {G}uido {G}overnatori, {H}o-{P}un {L}am, {A}ntonino {R}otolo, {S}erena {V}illata, \underline{\textbf{{G}hislain {A}temezing}} and {F}abien {G}andon: \textbf{{C}hecking licenses compatibility between vocabularies and data}. In 5th {I}nternational {W}orkshop on {C}onsuming {L}inked {D}ata (COLD 2014), {O}ctober 2014, {R}iva del {G}arda, {I}taly.

 \item \underline{\textbf{{G}hislain {A}temezing}}, {N}athalie {A}badie, {R}apha{\"e}l {T}roncy and {B}{\'e}n{\'e}dicte {B}ucher: \textbf{{P}ublishing Reference Geodata on the Web: {O}pportunities and Challenges for {IGN} {F}rance}. In 6th {I}nternational {W}orkshop on the {F}oundations, {T}echnologies and {A}pplications of the {G}eospatial {W}eb (TERRA COGNITA 2014), {O}ctober 2014, {R}iva del {G}arda, {I}taly.

 \item {G}uido {G}overnatori, {H}o-{P}un {L}am, {A}ntonino {R}otolo, {S}erena {V}illata, \underline{\textbf{{G}hislain {A}temezing}} and {F}abien {G}andon: \textbf{LIVE: a Tool for Checking Licenses Compatibility between Vocabularies and Data}. In 13th International Semantic Web Conference (ISWC 2014), Demo Track, {O}ctober 2014, {R}iva del {G}arda, {I}taly.

 \item \underline{\textbf{Ghislain Atemezing}} and Rapha{\"e}l Troncy: \textbf{{I}nformation content based ranking metric for linked open vocabularies}. In 10th {I}nternational {C}onference on {S}emantic {S}ystems (SEMANTICS 2014), {S}eptember 2014, {L}eipzig, {G}ermany.

 \item {A}hmad {A}ssaf, \underline{\textbf{Ghislain Atemezing}}, {R}apha{\"e}l {T}roncy and {E}lena {C}abrio: \textbf{{W}hat are the important properties of an entity? {C}omparing users and knowledge graph point of view}. In 11th {E}xtended {S}emantic {W}eb {C}onference (ESWC 2014), Demo Track, May 2014, {H}eraklion, {C}rete.

 \item Rapha{\"e}l Troncy, \underline{\textbf{Ghislain Atemezing}}, Nathalie Abadie and Cao-Vien Lam: \textbf{Modeling geometry and reference systems on the web of data}. In W3C Workshop on Linking Geospatial Data, March 2014, London, UK.

 \item \underline{\textbf{Ghislain Atemezing}}, Bernard Vatant, Rapha{\"e}l Troncy and Pierre-Yves Vanderbussche: \textbf{Harmonizing services for LOD vocabularies: a case study}. In Workshop on Semantic Web Enterprise Adoption and Best Practice (WASABI 2013), October, 2013, Sydney, Australia.

 \item Abdelfettah Feliachi, Nathalie Abadie, Hamdi Fay\c cal and \underline{\textbf{Ghislain Atemezing}}: \textbf{{I}nterlinking and visualizing linked open data with geospatial reference data}. In 9th International Workshop on Ontology Matching (OM 2014), October 2013, Sydney, Australia.

 \item \underline{\textbf{Ghislain Atemezing}}, Fabien Gandon, Gabriel Kepeklian, Fran\c{c}ois Scharffe, Rapha{\"e}l Troncy and Serena Villata: \textbf{When publishing linked data requires more than just using a tool}. In W3C Workshop on Open Data on the Web, April 2013, London, UK.

 \item \underline{\textbf{Ghislain Atemezing}} and Rapha{\"e}l Troncy: \textbf{Towards interoperable visualization applications over linked data}. In 2nd European Data Forum (EDF 2013), April 2013, Dublin, Ireland.

 \item Fran\c cois Scharffe, \underline{\textbf{Ghislain Atemezing}}, Rapha\"{e}l Troncy, Fabien Gandon, Serena Villata, B\'{e}n\'{e}dicte Bucher, Fay\c cal Hamdi, Laurent Bihanic, Gabriel K\'{e}p\'{e}klian, Franck Cotton, J\'{e}r\^{o}me Euzenat, Zhengjie Fan, Pierre-Yves Vandenbussche and Bernard Vatant: \textbf{Enabling linked-data publication with the datalift platform}. In AAAI 26th Conference on Artificial Intelligence, W10:Semantic Cities, July 2012, Toronto, Canada.

 \item \underline{\textbf{Ghislain Atemezing}} and Rapha\"{e}l Troncy: \textbf{Vers une meilleure interop\'{e}rabilit\'{e} des donn\'{e}es g\'{e}ographiques fran\c caises sur le Web de donn\'{e}es}. In 23\'{e}mes Journ\'{e}es Francophones d'Ing\'{e}nierie des Connaissances (IC 2012), June 2012, Paris, France.

 \item Houda Khrouf, \underline{\textbf{Ghislain Atemezing}}, Giuseppe Rizzo, Rapha\"{e}l Troncy and Thomas Steiner: \textbf{Aggregating social media for enhancing conference experiences}. In 1st International Workshop on Real-Time Analysis and Mining of Social Streams (RAMSS 2012), June 2012, Dublin, Ireland.

 \item Houda Khrouf, \underline{\textbf{Ghislain Atemezing}}, Thomas Steiner, Giuseppe Rizzo and Rapha\"{e}:l Troncy: \textbf{Confomaton: A conference enhancer with social media from the cloud}. In 9th Extended Semantic Web Conference (ESWC 2012), Demo Track, May 2012, Heraklion, Crete.

\end{enumerate}

\section*{W3C Documents}
\label{sec:w3cdocs}
\begin{enumerate}
 \item {B}ernadette {H}yland, \underline{\textbf{{G}hislain} {A}temezing} and {B}oris {V}illaz{\'o}n-{T}errazas: \textbf{Best Practices for Publishing Linked Data}. W3C Working Group Note published on January 9, 2014. URL: \url{http://www.w3.org/TR/ld-bp/}

 \item {B}ernadette {H}yland, \underline{\textbf{{A}temezing, {G}hislain}}, {M}ichael {P}endleton, {B}iplav {S}rivastava: \textbf{Linked Data Glossary}. W3C Working Group Note published on June 27, 2013. URL: \url{www.w3.org/TR/ld-glossary/}
\end{enumerate} 